The Ranked Nearest Neighbour Interchange space

### 2019-05-26 paper.tex
few details in proof of caterpillar convexity

    Induction basis: $l = \Delta(\rnniu) = \frac{(n-1)(n-2)}{2}$\\
    With Lemma~\ref{lemma:caterpillar_dist=diameter} one can see that for each pair of caterpillar trees $T$, $T'$ with distance $d(T,T') = \Delta(\rnniu)$ there is a caterpillar path that is a shortest path from $T$ to $T'$.

    Induction step: $l+1 \to l$\\
    Induction hypothesis: For each pair of caterpillar trees with distance $l+1$ there is a shortest path between these trees that only consists of caterpillar trees.
    Let $T$ and $T'$ be caterpillar trees with caterpillar distance $l$.
    Since $l < \Delta(\rnniu) = \frac{(n-1)(n-2)}{2}$, there is one loop $k$ in the Bubble Sort algorithm from $T'$ to $T$ where there are less than $n-2-k$ swaps of neighboured taxa.
    This means that there is a taxon $a$ that has parent with rank $n-2-k$ in $T'$ and $a$ does not need to move all the way up starting at the cherry in loop $k$.
    Therefore, there must be a taxon $b$ which has parent with rank one less than the parent of $a$ in the tree at the beginning of loop number $k$.
    If we assume that loop $k$ is the first loop where a single taxon is not moved by $n-k-2$ $\rnni$ moves, it follows that $a$ and $b$ must be neighboured taxa in $T'$.
    Let $T''$ be the caterpillar tree resulting from $T'$ by exchanging $a$ and $b$.
    It follows that $T'$ has caterpillar distance $d_c(T,T'') = d_c(T,T') + 1 = l+1$ to $T$ and $d(T',T'') = 1$.
    We can apply the induction hypothesis on $T$ and $T''$ and know that $d(T,T'') = l+1$.
    If it was $d(T,T') < l$, it would hold $d(T,T'') \leq d(T,T') + d(T',T'') < l + 1$ which contradicts the induction hypothesis.
    We can conclude that $d(T,T') = l$.

>>>>

    Induction basis: $l = \Delta(\rnniu) = \frac{(n-1)(n-2)}{2}$\\
    With Lemma~\ref{lemma:caterpillar_dist=diameter} one can see that for each pair of caterpillar trees $T$, $T'$ with caterpillar distance $d_c(T,T') = \Delta(\rnniu)$, it is $d(T,T') = d_c(T,T')$, which means that there is a shortest path between $T$ and $T'$ that is a caterpillar path.

    Induction step: $l+1 \to l$\\
    Induction hypothesis: For each pair of caterpillar trees with caterpillar distance $l+1$ there is a shortest path between these trees that only consists of caterpillar trees, which means that the distance between these trees is $l+1$.
    Let $T$ and $T'$ be caterpillar trees with caterpillar distance $l$.
    Since $l < \Delta(\rnniu) = \frac{(n-1)(n-2)}{2}$, there is one loop $k$ in the Bubble Sort algorithm from $T'$ to $T$ where there are less than $n-2-k$ swaps of neighboured taxa.
    This means that there is a taxon $a$ that has parent with rank $n-2-k$ in $T$ and that does not need to move all the way up starting at the cherry in loop $k$.
    Therefore, there must be a taxon $b$ which has parent with rank one less than the parent of $a$ in the tree at the beginning of loop number $k$.
    If we assume that loop $k$ is the first loop where a single taxon is not moved by $n-k-2$ $\rnni$ moves, it follows that $a$ and $b$ must be neighboured taxa in $T'$.
    Let $T''$ be the caterpillar tree resulting from $T'$ by exchanging $a$ and $b$.
    It follows that $T'$ has caterpillar distance $d_c(T,T'') = d_c(T,T') + 1 = l+1$ to $T$ and $d(T',T'') = 1$.
    We can apply the induction hypothesis on $T$ and $T''$ and know that $d(T,T'') = l+1$.
    If it was $d(T,T') < l$, it would hold $d(T,T'') \leq d(T,T') + d(T',T'') < l + 1$ which contradicts the induction hypothesis.
    We can conclude that $d(T,T') = l$.
